### 2015-02-06 cav-2015-tex/abstract.tex

A fundamental challenge of using symbolic execution for software
analysis is the treatment of dynamically allocated data. Existing
static analysis techniques over-approximate the potential heaps,
dynamic symbolic execution techniques often under-approximate the
heaps, while generalized symbolic execution leads to excessive
case-splitting. This paper introduces a compact heap representation
along with heap update rules that (i) allow reads and writes of
symbolic references, (ii) does not require constraint rewriting, and
(iii) supports arbitrary recursive data structures. As such, the
representation and update rules capture all possible heaps that follow
a common execution path. More importantly, in contrast to
state-of-the-art techniques, the definition and update rules are
unconstrained and not subject to bounds such as the number of heap
locations, aliasing, or memory partitions. An evaluation of a
proof-of-concept implementation in the Java Pathfinder framework is
presented to demonstrate the computational feasibility of the approach
over several classical symbolic execution benchmarks.

>>>>

A fundamental challenge of using symbolic execution for software
analysis is the treatment of dynamically allocated data. Existing
static analysis techniques over-approximate the potential heaps,
dynamic symbolic execution techniques often under-approximate the
heaps, while generalized symbolic execution leads to excessive
case-splitting. This paper introduces a compact heap representation,
along with heap update rules that (i) allow reads and writes of
symbolic references, (ii) do not require constraint rewriting, and
(iii) support arbitrary recursive data structures. As such, the
representation and update rules capture all possible heaps that follow
a common execution path. More importantly, in contrast to existing
state-of-the-art techniques, the definition and update rules are
unconstrained and not subject to bounds, such as the number of heap
locations, aliasing, or memory partitions. An evaluation of a
proof-of-concept implementation in the Java Pathfinder framework is
presented to demonstrate the computational feasibility of the approach
over several classical symbolic execution benchmarks.

### 2015-02-06 cav-2015-tex/approach.tex
edits to the approach section

\subsection{Generating Symbolic Heaps}
\label{sec:precise}

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location.  In the field-access rewrite rule
in~\figref{fig:fHeap} gets the field type $\cfgnt{C}$ and uses
\figref{fig:symInit}, to get a new symbolic heap that is initialized
on the field $\cfgnt{f}$. The symbolic heap is further modified to
include a new stack reference pointing to the value set from the read;
the new stack reference is the return value from the operation.

@@

The value set computation is local: it only considers constraint
locations on the reference $\cfgnt{r}$. Further, it only includes
constraints that are feasible under the current path condition
$\phi_g$. It also ensures that the access path is valid. There is a
constraint $\phi$ from the base reference $r$ and a given
location. Once a location $l$ is selected its field $\cfgnt{f}$ is
accessed to get another reference $r^\prime$, and from that reference
there is a set of locations each with a constraint. The first
constraint $\phi$ from the base reference plus this second constraint
to the actual location of the field $\phi^\prime$ comprise the access
path. For each member of the value set, these constraints together
with the global path constraint must be satisfied on some aliasing
assignment. \nsr{neha: this para needs work -- flow and content}

For field-write in~\figref{fig:fHeap}, after the non-null check and
strengthening of the global heap constraint, it computes the current
references associated with the field in every location in
$\Psi_\cfgnt{x}$. Please note that the reference $\cfgnt{r}_x$ is the
base reference whose field, $\cfgnt{r}_\mathit{cur}$ is being written
to while the reference $r$ is the target reference. The set $\Psi_x$
contains tuples $(\phi\ l\ \cfgnt{r}_\mathit{cur})$ of constraints,
locations, and references. These tuples represent access chains
leading from $\cfgnt{r}_x$ to the reference of the field,
$\cfgnt{r}_\mathit{cur}$. The goal is to change the fields to no
longer point to $\cfgnt{r}_\mathit{cur}$, but rather fresh references
that have both the original locations before the write and the
locations from the write in the value sets (i.e., conditional
write). Since the target of the write is $r$, the rule checks that the
location constraint pairs of $\cfgnt{L}(\cfgnt{r})$ are satisfiable
when accessed through the $\cfgnt{r}_x$ chain in the strengthening
function.  \input{st-def} Constraints on locations are strengthened
with new aliasing conditions and those that are feasible with the
current path constraint are retained.

Strengthening in the field write creates a value set that contains two
types of locations: those for the case where the write is feasible
(the first call), and those for the case where it is not (the second
call).  This is the conditional in the write. In the case that $\phi$
is true then $\cfgnt{r}_\mathit{cur}$ will point to the constraint
location pairs of $\cfgnt{L}(\cfgnt{r})$ while if $\phi$ is false then
$\cfgnt{r}_\mathit{cur}$ will continue to point to the constraint
location pair it is currently referenced.  As the references are
immutable, the rule creates fresh references for each
$\cfgnt{r}_\mathit{cur}$ and points them to the appropriate value
sets.

@@

The rewrite rules to compare two references in the symbolic heap are
shown in~\figref{fig:eqs}. The equals references-true rewrite rule
returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$
\emph{can} be equal. Intutively, $\Phi_\alpha$ contains all
constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may point to
the same location in the symbolic heap. The second set, $\Phi_0$,
contains constraints under which the reference $\cfgnt{r}_0$ points to
corresponding locations such that the reference $\cfgnt{r}_1$
\emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and
$\cfgnt{r}_0$ \emph{does not} point to those locations under any
constraint. We use the three sets of constraints to update the current
global heap constraint $\phi_g$ and create a new global heap
constraint $\phi_g^\prime$. We add to $\phi_g$ the disjunction of the
constraints in $\Phi_\alpha$ to indicate that if any of the
constraints are satisfiable, then references $\cfgnt{r}_0$ and
$\cfgnt{r}_1$ can be equal. Furthermore, we conjoin to the global heap
constraint $\phi_g^\prime$, the conjunctions of negations to the
constraints in $\Phi_0$ and $\Phi_1$. This indicates for locations
that are not common to the references, the negations of their
constraints are satisfiable. The rule does not complete (i.e., is not
feasible) if the new global constraint is not satisfied on any
aliasing assignment. In such a case, the $\cfgt{true}$ outcome is not
possible on any symbolic heap. Before the rewrite rule returns
$\cfgt{true}$ we check the satisfiability of the updated global heap
constraint. The reference-false is the logical dual of the rule.

>>>>

\subsection{Generating and Updating Symbolic Heaps}
\label{sec:precise}

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location. The field-access rewrite rule
in~\figref{fig:fHeap} dereferences a field of type $\cfgnt{C}$ and
uses \figref{fig:symInit}, to get a new symbolic heap that is
initialized on the field $\cfgnt{f}$. The symbolic heap is further
modified to include a new stack reference pointing to the value set
(possible values of the field) returned during the dereferencing; the
new stack reference is the return value from the operation.

@@

The value set computation is local: it only considers constraint
locations on the reference $\cfgnt{r}$. Further, it only includes
constraints that are feasible under the current path condition
$\phi_g$. It also ensures that the access path is valid. There is a
constraint $\phi$ from the base reference $r$ and a given
location. Once a location $l$ is selected its field $\cfgnt{f}$ is
accessed to get another reference $r^\prime$, and from that reference
there is a set of locations each with a constraint. The first
constraint $\phi$ from the base reference plus this second constraint
to the actual location of the field $\phi^\prime$ comprise the access
path. For each member of the value set, these constraints together
with the global path constraint must be satisfied on some aliasing
assignment. % \nsr{neha: this para needs work -- flow and content}

For field-write in~\figref{fig:fHeap}, after the non-null check and
strengthening of the global heap constraint, it computes the current
references associated with the field in every location in
$\Psi_\cfgnt{x}$. Note that the reference $\cfgnt{r}_x$ is the base
reference whose field, $\cfgnt{r}_\mathit{cur}$ is being written to,
while the reference $r$ is the target reference. The set $\Psi_x$
contains tuples $(\phi\ l\ \cfgnt{r}_\mathit{cur})$ of constraints,
locations, and references. These tuples represent access chains
leading from $\cfgnt{r}_x$ to the reference of the field,
$\cfgnt{r}_\mathit{cur}$. The goal is to change the fields to no
longer point to $\cfgnt{r}_\mathit{cur}$, but rather fresh references
that have both the original locations before the write and the
locations from the write in the value sets (i.e., conditional
write). Since the target of the write is $r$, the rule checks that the
location constraint pairs of $\cfgnt{L}(\cfgnt{r})$ are satisfiable
when accessed through the $\cfgnt{r}_x$ chain in the strengthening
function.  \input{st-def} Constraints on locations are strengthened
with new aliasing conditions and those that are feasible with the
current path constraint are retained.

Strengthening in the field write creates a value set that contains two
types of locations: those for the case where the write is feasible
(the first call), and those for the case where it is not (the second
call). In the case that $\phi$ is true then $\cfgnt{r}_\mathit{cur}$
will point to the constraint location pairs of $\cfgnt{L}(\cfgnt{r})$.
Whereas, if $\phi$ is false then $\cfgnt{r}_\mathit{cur}$ will
continue to point to the constraint location pair it currently
references.  As the references are immutable, the rule creates fresh
references for each $\cfgnt{r}_\mathit{cur}$ and points them to the
appropriate value sets.

@@

The rewrite rule to compare two references in the symbolic heap is
shown in~\figref{fig:eqs}. The equals references-true rewrite rule
returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$
\emph{can} be equal. Intutively, $\Phi_\alpha$ contains all
constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may point to
the same location in the symbolic heap. The second set, $\Phi_0$,
contains constraints under which the reference $\cfgnt{r}_0$ points to
corresponding locations such that the reference $\cfgnt{r}_1$
\emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and
$\cfgnt{r}_0$ \emph{does not} point to those locations under any
constraint. We use the three sets of constraints to update the current
global heap constraint $\phi_g$ and create a new global heap
constraint $\phi_g^\prime$. We add to $\phi_g$ the disjunction of the
constraints in $\Phi_\alpha$ to indicate that if any of the
constraints are satisfiable, then references $\cfgnt{r}_0$ and
$\cfgnt{r}_1$ can be equal. Furthermore, we conjoin to the global heap
constraint $\phi_g^\prime$, the conjunctions of negations to the
constraints in $\Phi_0$ and $\Phi_1$. This indicates for locations
that are not common to the references, the negations of their
constraints are satisfiable. The rule does not complete (i.e., is not
feasible) if the new global constraint is not satisfied on any
aliasing assignment. In such a case, the $\cfgt{true}$ outcome is not
possible on any symbolic heap. Before the rewrite rule returns
$\cfgt{true}$ we check the satisfiability of the updated global heap
constraint. The reference-false is the logical dual of the rule.

### 2015-02-06 cav-2015-tex/summary-heap.tex

Our symbolic heap is a field sensitive points-to graph with edges
annotated with disjunctive formulas describing aliasing
relationships. We represent the symbolic heap is a bipartite graph on
references, $\cfgnt{r}$, and constraint location pairs,
$\lp\phi\ \cfgnt{l}\rp$. The graph is expressed in the location map,
$\cfgnt{L}$, and the reference map, $\cfgnt{R}$. As done with the
environment, $\cfgnt{L}$ and $\cfgnt{R}$ are treated as partial
functions where $\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the set
of constraint-location pairs in the heap associated with the given
reference, and $\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the
reference associated with the given location-field pair in the
heap. References in the graph are immutable. Locations, however, are
mutated on updates.  Immutability of references means that look-ups in
the heap are non-recursive and updates do not modify the current path
condition.

The soundness and completeness proofs \sjp{in Section...} rely on a set of 
invariants \sjp{properties of the symbolic heap?} that
are preserved by the small-step semantics during symbolic execution:
%\emph{reference partitions}, \emph{null and uninitialized}, \emph{immutability}, \emph{determinism}, and
%\emph{type consistency}.
\begin{compactdesc}
\item[Reference partitions] means that a reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap} that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in
  constraints); $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals} that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere); and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} (these appear in environments, expressions, and continuations).
\item[Null and Uninitialized] means that every symbolic heap contains 
a special location for null ($\cfgnt{l}_\mathit{null}$), and a special
location for uninitialized
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability] means that a reference does not change in the
location map, and as a consequence, constraints never need to be
rewritten as the program evolves its state.
\item[Determinism] means that a reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency] means that all locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

%\sjp{Up to now, no discussion of lazy initialization - what it is or
%  how it is being used.}  \nsr{A key insight to the summary heap is
%  that although many auxiliary or stack literals are created during
%  symbolic execution with lazy initialization, only input heap
%  references appear in the constraints expressing potential aliasing
%  relationships.}
A solver is used to reason over the aliasing relationships, and a
satisfying assignment uniquely identifies the set of aliasing
relationships that are representative of a concrete heap and its
shape. For example, if the solver assigns input heap variables
$\cfgnt{r}$ and $\cfgnt{r}^\prime$ such that $\cfgnt{r} =
\cfgnt{r}^\prime$ is true, then there exists some location $\cfgnt{l}$
and constraints $\phi$ and $\phi^\prime$ such that (i)
$\lp\phi\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r})$; (ii)
$\lp\phi^\prime\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r}^\prime)$; and
(iii) $\phi$ and $\phi^\prime$ hold in the aliasing assignments from
the solver.  In this way, the graph groups heaps with the conditions
under which those heaps exist so look-ups are non-recursive and
updates do not modify the path condition or require rewrites on
constraints.

>>>>

Our symbolic heap is a field sensitive points-to graph with edges
annotated with disjunctive formulas describing aliasing
relationships. We represent the symbolic heap as a bipartite graph on
references, $\cfgnt{r}$, and constraint location pairs,
$\lp\phi\ \cfgnt{l}\rp$. The graph is expressed in the location map,
$\cfgnt{L}$, and the reference map, $\cfgnt{R}$. As done with the
environment, $\cfgnt{L}$ and $\cfgnt{R}$ are treated as partial
functions where $\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the set
of constraint-location pairs in the heap associated with the given
reference, and $\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the
reference associated with the given location-field pair in the
heap. References in the graph are immutable. Locations, however, are
mutated on updates.  Immutability of references means that look-ups in
the heap are non-recursive and updates do not modify the current path
condition.

The soundness and completeness proofs \sjp{in Section...} rely on a set of 
invariants \sjp{properties of the symbolic heap?} that
are preserved by the small-step semantics during symbolic execution:
%\emph{reference partitions}, \emph{null and uninitialized}, \emph{immutability}, \emph{determinism}, and
%\emph{type consistency}.
\begin{compactdesc}
%<<<<<<< HEAD
\item[Reference partitions:] Each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}% that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in constraints)
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}% that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere)
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} %(these appear in environments, expressions, and continuations).
\item[Null and Uninitialized:] Every summary heap contains 
two special locations, null ($\cfgnt{l}_\mathit{null}$), and uninitialized,
%=======
%\item[Reference partitions] means that a reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
%  heap} that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in
%  constraints); $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
%  literals} that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere); and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
%  literals} (these appear in environments, expressions, and continuations).
%\item[Null and Uninitialized] means that every symbolic heap contains 
%a special location for null ($\cfgnt{l}_\mathit{null}$), and a special
%location for uninitialized
%>>>>>>> 49c83f34a0b5afd58adc4b6649b15e2cbc93437c
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability:] References do not change once added to the
location map, eliminating any need for constraint rewriting as the program state evolves.
\item[Determinism:] A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency:] All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

%\sjp{Up to now, no discussion of lazy initialization - what it is or
%  how it is being used.}  \nsr{A key insight to the summary heap is
%  that although many auxiliary or stack literals are created during
%  symbolic execution with lazy initialization, only input heap
%  references appear in the constraints expressing potential aliasing
%  relationships.}
A solver is used to reason over the aliasing relationships, and a
satisfying assignment uniquely identifies the set of aliasing
relationships that are representative of a concrete heap and its
shape. For example, if the solver assigns input heap variables
$\cfgnt{r}$ and $\cfgnt{r}^\prime$ such that $\cfgnt{r} =
\cfgnt{r}^\prime$ is true, then there exists some location $\cfgnt{l}$
and constraints $\phi$ and $\phi^\prime$ such that (i)
$\lp\phi\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r})$; (ii)
>>>>>>> 49c83f34a0b5afd58adc4b6649b15e2cbc93437c
$\lp\phi^\prime\ \cfgnt{l}\rp \in \cfgnt{L}(\cfgnt{r}^\prime)$; and
(iii) $\phi$ and $\phi^\prime$ hold in the aliasing assignments from
the solver.  In this way, the graph groups heaps with the conditions
under which those heaps exist so look-ups are non-recursive and
updates do not modify the path condition or require rewrites on
constraints.
